---
title: "Model Fitting"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
  always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)

stratified = 
  read_csv("https://raw.githubusercontent.com/Serena-Wang/p8105_covid_vaccinations/main/Data/coverage_by_demo_full.csv") %>%
  janitor::clean_names() %>%
  select(group, subgroup, perc_partially, perc_fully, perc_additional) %>%
  drop_na()

unstratified = 
  read_csv("https://raw.githubusercontent.com/Serena-Wang/p8105_covid_vaccinations/main/Data/coverage_boro_demo_full.csv") %>%
  janitor::clean_names() %>%
  select(age_group, race_ethnicity, city_perc_fully) %>%
  filter(
    !age_group %in% c("Boroughwide", "All ages", "65+", "13-17", "18-44", "45-64") 
)
```

## Model Fitting

We propose 3 models for prediction based on previous analysis: 

 1. Linear Model of city_perc_fully ~ age_group  

 2. Linear Model of city_perc_fully ~ age_group  +  race_ethnicity

 3. Linear Model of city_perc_fully ~ age_group  +  race_ethnicity + age_group * race_ethnicity, assuming that there are interaction

### Age Group Model

Age group model is a simple linear model that contains only the most significant predictor: age of the NYC residents. Age predictor has five levels: 5-12, 13-17, 18-44, 45-64 and 65+.
```{r}
fit_lm_age = lm(city_perc_fully ~ age_group, data = unstratified  )
fit_lm_age  %>% broom::tidy() %>% knitr::kable(digits = 2)
```

### Main Effect Model

Main effect model contains is a multiple linear regression model with two predictors: age and race/ethnicity. Age predictor has five levels: 5-12, 13-17, 18-44, 45-64 and 65+. And race group has three levels: Black, White, and Hispanic/Latino.
```{r}
fit_lm_main = lm(city_perc_fully ~ age_group  +  race_ethnicity, 
                 data = unstratified )
fit_lm_main  %>% broom::tidy() %>% knitr::kable(digits = 2)
```

### Interaction Model

Interaction model is a multiple linear regression model with three predictiors: age, race, and age*race. 
```{r}
fit_lm_interaction = lm(city_perc_fully ~ age_group  +  race_ethnicity +
                        age_group * race_ethnicity, 
                        data = unstratified )
fit_lm_interaction  %>% broom::tidy() %>% knitr::kable(digits = 2)
```

## Cross Validation 

Although race is not statistically significant under the critical value alpha = 0.05, it increases the prediction ability of the simple linear model : lm(city_perc_fully ~ age_group). What's more, we notice that interaction term is contributing to model predictability as well. We would say the interaction model  will be better than the other one because its root means square error seems to be smaller. Therefore, interaction model could be useful for predicting the fully vaccinated rate in NYC.

```{r}
cv_df =
  crossv_mc(unstratified, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(
    age_mod = map(train, ~lm(city_perc_fully ~ age_group, data = .x)),
    main_mod  = map(train, ~lm(city_perc_fully ~ age_group  +  race_ethnicity, data = .x)),
    interaction_mod = map(train, ~lm(city_perc_fully ~ age_group  +  race_ethnicity + age_group * race_ethnicity, data = .x))) %>% 
  mutate(
    rmse_age = map2_dbl(age_mod, test, ~rmse(model = .x, data = .y)),
    rmse_main = map2_dbl(main_mod, test, ~rmse(model = .x, data = .y)),
    rmse_interaction = map2_dbl(interaction_mod, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

## Model Conclusion

We select the linear model of city_perc_fully ~ age_group  +  race_ethnicity + age_group * race_ethnicity

Age, race and their interaction term are good predictors for anticipating the COVID vaccination rate in NYC.

